---
title: "First Week Quiz"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)

```


## Definitions & Concepts

#### Select one correct answer

```{r q1}
  question("What is the definition of Data Science?",
    answer("Multi-disciplinary field that uses scientific methods and algorithms to extract knowledge", correct = TRUE),
    answer("Multi-disciplinary field that uses statistical methods to predict the target values"),
    answer("It is a computer science field that focus on algorithms and models to gain insight"),
    answer("Data science is not actually a science"), 
    allow_retry = TRUE, 
    random_answer_order = TRUE
 
)
```

```{r q2}
  question("Which order is the true order of level of measurements of data? (from lowest level to highest level)",
    answer("Ordinal, Ratio, Catagorical, Nominal"),
    answer("Nominal, Ordinal, Interval, Ratio", correct = TRUE),
    answer("Nominal, Ordinal, Ratio, Interval" ),
    answer("Ratio, Nominal, Ordinal, Interval"),
    answer("Ordinal, Ratio, Nominal, Catagorical"),
    allow_retry = TRUE, 
    random_answer_order = TRUE
)
```


```{r q3}
  question("In CRISP-DB, Business Understanding phase is about",
    answer("Identifying the problem and setting the objective of the project", correct = TRUE),
    answer("Understand everything about the business"),
    answer("Collecting data for analysis"),
    answer("Opening new business"),
    allow_retry = TRUE, 
    random_answer_order = TRUE
)
```

```{r q4}
  question("In CRISP-DB, Data Preparation phase is about",
    answer("Cleaning up the data and preprocesing it into suitable format for modeling", correct = TRUE),
    answer("Understanding the way the data was collected and the systems generating data"),
    answer("Saving data in excel format"),
    answer("Making sure computer hard-drive is clean from dust"),
    allow_retry = TRUE, 
    random_answer_order = TRUE
)
```

```{r q5}
  question("The process of breaking tables apart and thereby reducing data redundancy is called",
    answer("Normalization", correct = TRUE),
    answer("Duplication"),
    answer("Transformation"),
    answer("Table Breaking"),
    allow_retry = TRUE, 
    random_answer_order = TRUE
)
```


#### Select all correct answers. 

```{r q6}
  question("Select all that are examples of structured data",
    answer("Family photos"),
    answer("Phone contacts", correct = TRUE),
    answer("Customers Records", correct = TRUE),
    answer("Blogs"),
    answer("Medical Appoinments", correct = TRUE),
    answer("Youtube Videos"),
    allow_retry = TRUE, 
    random_answer_order = TRUE,
    type = "multiple"
)
```


```{r q7}
  question("Select all that are examples of unstructured data",
    answer("presentations", correct = TRUE),
    answer("Phone contacts"),
    answer("Customers Records"),
    answer("Songs", correct = TRUE),
    answer("Medical Appoinments"),
    answer("Youtube Videos", correct = TRUE),
    allow_retry = TRUE, 
    random_answer_order = TRUE,
    type = "multiple"
)
```

```{r q8}
  question("Select all that considered main tasks of data scientist",
    answer("Describe data", correct = TRUE),
    answer("Create new data"),
    answer("Classify data", correct = TRUE),
    answer("Predict data", correct = TRUE),
    answer("Buy data"),
    answer("Identify associations with data", correct = TRUE),
    allow_retry = TRUE, 
    random_answer_order = TRUE,
    type = "multiple"
)
```


#### True or False

```{r q9}
  question("Nominal and Ordinal are catagorical types",
    answer("TRUE", correct = TRUE),
    answer("FALSE"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```


```{r q10}
  question("Ratio level has no absolute zero",
    answer("TRUE"),
    answer("FALSE", correct = TRUE),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```


## Git commands

#### Select one correct answer

```{r q11}
  question("In order to initiate new Git repository we use the comand",
    answer("`git init`", correct = TRUE),
    answer("`git status`"),
    answer("`git branch`"),
    answer("`git commit`"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```

```{r q12}
  question("You created and new file named `fun.txt` in your working directory, you made changes and you want to commit, how do you do it?",
    answer("Start tracking the file first by `git add fun.txt` then commit", correct = TRUE),
    answer("Start by using `git status` to see the current changes then commit"),
    answer("Git will add the file automatically when you run this command `git commit`"),
    answer("You can't add new files to the Git repository after initiation"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```

```{r q13}
  question("How to tell Git to ignore certain files?",
    answer("Add the file name to `.gitignore` file", correct = TRUE),
    answer("You need to create sub-repository that parent git can't track"),
    answer("Change the name of the file everytime you make changes"),
    answer("Simply `git ignore` then file name"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```

```{r q14}
  question("How to merge to a branch `featureA` with branch `featureB`",
    answer("we checkout to `featureA` by `git checkout feastureA` then `git merge feastureB`", correct = TRUE),
    answer("we checkout to `featureB` by `git checkout feastureB` then `git merge feastureA`"),
    answer("`git merge feastureA featureB`"),
    answer("`git merge`, Git will know what to do"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```

```{r q15}
  question("How to switch to a branch named `feastureB`",
    answer("`git branch featureA`", correct = TRUE),
    answer("`git checkout -b featureA`", correct = TRUE),
    answer("`git branch -d featureA`", message = "be careful this will actually delete the branch `featureA` if it exists"),
    answer("`git create branch featureA`"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```


#### Select all correct answers. 

```{r q16}
  question("How to start new branch named `featureA`",
    answer("`git branch featureA`", correct = TRUE),
    answer("`git checkout -b featureA`", correct = TRUE),
    answer("`git branch -d featureA`", message = "be careful this will actually delete the branch `featureA` if it exists"),
    answer("`git create branch featureA`"),
    random_answer_order = TRUE,
    allow_retry = TRUE,
    type = "multiple"
)
```


## R Programming

#### All the below are exercises that are not evaluated but good for practice 

##### Operations, Variables and Types.  

```{r preperations}
x <- list(a = 19, b = "Data Science", c = matrix(1:10, nrow = 5, byrow = TRUE))
y <- as.Date("2020-02-14")
df <- data.frame(numbers = sample(10), products = c("Bread", "Rice", "Olive Oil", "Soup", "Chocolate", "Milk", "Banana", "Cheese", "Soda", "Apple"))
cars_df <- cars
```
Perform the follwoing taks:

  - Add 13 to 347 then divide by 32.(_bonus_): round to nearest integer,_hint_: use `round()` functions. 


```{r q17, exercise = TRUE, exercise.lines = 3, exercise.eval = FALSE}

```

  - What is the class of the `x`?
  - What is the class of element `c` in `x`?
  - Is element `b` an integer? _bonus_: if character how many characters?
  - Change the value of element `a` to `100`
  - Multiple element `a` by `3.4` and save it in element `d`
  - What is the class of element `d`. _think why `d` is not integer_

```{r q18, exercise = TRUE, exercise.lines = 15,exercise.setup = "preperations", exercise.eval = FALSE}
x
```

##### Dates  

###### Challenge 

Install the ``lubridate`` package and refer to the [cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/lubridate.pdf) to perform the following 

- Extract the year from variable `y`
- Extract the name of the month from variable `y`
- What day of the week is `y`
- What would the date be after 14 days from `y`

```{r q19, exercise = TRUE, exercise.lines = 15,exercise.setup = "preperations", exercise.eval = FALSE}
#library("lubridate")
y
```


##### Data Frames  

Perform the following for the data frame `df` 

- Display the first 10 records from `df`
- Extract the dimension of the `df`. _hint_: get the number of rows and columns.
- Change the name of the first column to `quantity`
- Slice the `df` to include only the second column and rows from 4 to 7._hint_: to perserve type of data.farme use `drop` argument.
- Print summary of the data frame. 

```{r q20, exercise = TRUE, exercise.lines = 15,exercise.setup = "preperations", exercise.eval = FALSE}
df
```

##### Graphing in R

Perform the following in both base R and `ggplot2` package.

- Generate a histogram of the `speed` column with 10 bins. *bonus*: color graph with "violet" and set title to "Histogram" and x label to "Speed".
- Generate a scatterplot with `speed` on the y-axis and `dist` on x-axis. *bonus*: color values above 10 differently.  
- Generate a boxplot for `dist`

###### Base R 

- Generate a histogram of the `speed` column with 10 bins. *bonus*: color graph with "violet" and set title to "Histogram" and x label to "Speed".
- Generate a scatterplot with `speed` on the y-axis and `dist` on x-axis. *bonus*: color values above 10 differently.  
- Generate a boxplot for `dist`

```{r q21, exercise = TRUE, exercise.lines = 15,exercise.setup = "preperations", exercise.eval = FALSE}
cars_df
```

###### `ggplot2` graphics 

Use the `ggplot2` package and refer to the [cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf) to perform the previous graphs

```{r q22, exercise = TRUE, exercise.lines = 20,exercise.setup = "preperations", exercise.eval = FALSE}
cars_df
```



```{r q17-solution}
(13+347)/32
```

```{r q18-solution}
#Class of x
class(x)
#class of element c in x
class(x$c)
#checking if element b is numeric 
is.integer(x$b)
nchar(x$b)
#Changing the value of element a to 100
x$a <- 100
#Multiple element `a` by `3.4` and save it in element `d`
x$d <- x$a * 3.4
#What is the class of element `d`
class(x$d)
```

```{r q19-solution}
library("lubridate")
#extracting the year
year(y)
#extracting the name of the month
month(y,label = TRUE, abbr = FALSE)
#Day of the week is 
wday(y, label = TRUE, abbr = FALSE)
#What would the date be after 14 days from `y`
y + days(14)
```


```{r q20-solution}
#Display the first 6 records from `df`
head(df,n = 6)
#Extract the dimension of the `df`. _hint_: get the number of rows and columns.
dim(df)
#Change the name of the first column to `quantity`
colnames(df)[1] <- "quantity"
#Slice the `df` to include only the second column and rows from 4 to 7. 
df[4:7,2, drop = FALSE]
#Print summary of the data frame. 
summary(df)
```

```{r q21-solution}
#Generate a histogram of the `speed` column
hist(cars_df$speed, breaks = 10)
# Bonus question
hist(cars_df$speed, breaks = 10,col = "violet", main = "Histogram", xlab = "Speed")
#Generate a scatterplot with `speed` on the y-axis and `dist` on x-axis. *bonus*: color speed values above 10 differently. 
plot(y =cars_df$speed, x = cars_df$dist)
# Bonus question
plot(y =cars_df$speed, x = cars_df$dist, col = ifelse(cars_df$speed > 10, "black", "red"))
#Generate a boxplot for `dist`
boxplot(cars_df$dist)
```

```{r q22-solution}
library("ggplot2")
#Generate a histogram of the `speed` column
ggplot(data = cars_df, mapping = aes(x = speed)) + geom_histogram(bins = 10, color = "black")
# Bonus question
ggplot(data = cars_df, mapping = aes(x = speed)) + 
  geom_histogram(bins = 10, color = "black", fill ="violet") +
  labs(title = "Histogram", xlab = "Speed")
#Generate a scatterplot with `speed` on the y-axis and `dist` on x-axis. *bonus*: color speed values above 10 differently. 
ggplot(data = cars_df, mapping = aes(x = dist,y = speed)) +
  geom_point()
# Bonus question
ggplot(data = cars_df, mapping = aes(x = dist,y = speed)) +
  geom_point(aes(color =ifelse(speed > 10, "black", "red")),show.legend = FALSE)
#Generate a boxplot for `dist`
ggplot(data = cars_df, mapping = aes(y = dist)) + geom_boxplot()
```

