---
title: "Graduation Project"
author: "ENTER YOUR NAME HERE"
date: "" # mm/dd/yyyy
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load all the necessary libraries and comment why did you need them.

```

## Study Case: churning customers from a Telecom company. 

## Background Story 

Now that you have taken an amazing course with your nice teacher the Arabian Analyst you decided to open a data science consulting firm. Your first customer is a Telecom company that is trying to figure out the reasons why customers are switching to their competitors.  During your meeting with the CEO Ola, 

## Data Dictionary 

## Analysis 

### Setp 1: Importing the Data

```{r Importing the data}
# use read.csv or read_csv from readr packages 
```

### Setp 2: Split the dataset into training and testing sets

```{r}
# use createDataPartition() from caret package or use the sample() from base R. 
```



### Setp 3: Descriptive statistics

```{r}
# use summary() function or perform more advance descriptive statistics using dplyr package
```

Write you analysis from Descriptive statistics here .... 

#### BONUS Step 3.1 : Exploratory Analysis
```{r}
# use ggplot to explore :
  # data centrality 
  # data variability 
  # data shape 
  # outliers 
```

Write you analysis from Exploratory Analysis here .... 


##### Correlations 
```{r}
# Describe the relationship between variables 
```

Write you analysis from Correlations here .... 

*hint*: identify highly correlated variables and explain why we should not include them. 

##### Histograms

```{r}
# use ggplot package with geom_histogram() or hist() from base R. 
```


Write you analysis from Histograms here .... 
*hint*: identify skewed variables and determine if you need to do transformation on them or not. Justify your decision. 



##### Scatterplots 

```{r}
# use ggplot package with geom_point() or plot() from base R. 
```

Write you analysis from Scatterplots here .... 

#### Boxplots 

```{r}
# use ggplot package with geom_boxplot() or boxplot() from base R. 
```

Write you analysis from Boxplots here .... 

### Setp 5: Create dummy variables for categorical data


```{r}
# Use dummyVars() from caret package or perform the conversion manually. 
```

Describe which categorical variables did you convert to dummy variables and why you chose them. 


### Setp 6: Data Normalization and Transformation

```{r}
# Use perProcess() from caret package or perform the conversion manually. 
```

Explain why it is important to normalize the data. Moreover, based on what you decision after performing histograms, you should perform all necessary transformation here.


### Setp 7: Training and Tuning Model

```{r}
# use train() from caret package or select your chose from any one of the models we discussed during this course
```

justify why you chose the model you did and discuss ways that this model could be improved. 

*hint*: use cross validation and fine tune your model by using a grid search. 

### Setp 8: Testing and Evaluating the model


```{r}
# use the confusionMatrix() from caret package or confusion() form DAAG package. 
# make sure that the output data is in the right format for those functions. 
```


Discuss your results and explain why your model is performing well or not well. 


