---
title: "Third Week Quiz"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library("dplyr")
```

## Definitions & Concepts

#### Select one correct answer

```{r q1}
  question("Classification is?",
    answer("To assign a class based on based on predictor variables", correct = TRUE),
    answer("To estimate numerical value based on predictor variables"),
    answer("To group based on patterns between all variables"),
    answer("Not good practice in data science"), 
    allow_retry = TRUE, 
    random_answer_order = TRUE
 
)
```

```{r q2}
  question("Supervised learning is?",
    answer("modeling technique that is used when labeled data is available", correct = TRUE),
    answer("modeling technique that is used when labeled data is not available"),
    answer("modeling technique that tries to generate data based on human input"),
    answer("Not good practice in data science"), 
    allow_retry = TRUE, 
    random_answer_order = TRUE
 
)
```

```{r q3}
  question("Unpervised learning is?",
    answer("modeling technique that is used when labeled data is available"),
    answer("modeling technique that is used when labeled data is not available", correct = TRUE),
    answer("modeling technique that tries to generate data based on human input"),
    answer("Not good practice in data science"), 
    allow_retry = TRUE, 
    random_answer_order = TRUE
 
)
```

```{r q4}
  question("We get the *Sum of Squares Total Errors (SST)* by calculating ..",
    answer("sum of the squared difference between the dependant variable and the mean", correct = TRUE),
    answer("sum of the difference between the dependant variable and the mean"),
    answer("difference between the dependant variable and the mean"),
    answer("sum of the squared difference between the dependant variable and predicted value"), 
    allow_retry = TRUE, 
    random_answer_order = TRUE
 
)
```


```{r q5}
  question("We get the *Sum of Squares Errors (SSE)* by calculating ..",
    answer("sum of the squared difference between the dependant variable and the mean"),
    answer("sum of the difference between the dependant variable and the predicted value"),
    answer("difference between the dependant variable and the predicted value"),
    answer("sum of the squared difference between the dependant variable and predicted value", correct = TRUE), 
    allow_retry = TRUE, 
    random_answer_order = TRUE
 
)
```

```{r q6}
  question("We get the *Sum of Squares Regression (SSR)* by calculating ..",
    answer("sum of the squared difference between the predicted value variable and the mean", correct = TRUE),
    answer("sum of the difference between the dependant variable and the predicted value"),
    answer("sum of the difference between the dependant variable and the mean"),
    answer("sum of the squared difference between the dependant variable and predicted value"), 
    allow_retry = TRUE, 
    random_answer_order = TRUE
 
)
```

#### Select all correct answers. 

```{r q7}
  question("the term **MSE** is ",
    answer("$$\\frac{SSE}{n-2}$$", correct = TRUE),
    answer("Mean of Squared Error", correct = TRUE),
    answer("$$\\sum_{i=1}^{n}( \\hat{y}_i-\\bar{y})^2$$"),
    answer("The root of the mean squared error"), 
    allow_retry = TRUE, 
    random_answer_order = TRUE, 
    type = "multiple"
)
```


```{r q8}
  question("the term **RMSE** is ",
    answer("$$\\sqrt{\\frac{SSE}{n-2}}$$", correct = TRUE),
    answer("The mean squared error"),
    answer("$$\\sum_{i=1}^{n}(y_{i} - \\bar{y})^2$$"),
    answer("The root of the mean squared error", correct = TRUE), 
    allow_retry = TRUE, 
    random_answer_order = TRUE, 
    type = "multiple"
)
```

```{r q9}
  question("The **Coefficient of Determination** is the same as  ",
    answer("$$\\frac{SSR}{SST}$$", correct = TRUE),
    answer("$$R$$"),
    answer("$$y = a + bx + \\epsilon$$"),
    answer("$$R^2$$", correct = TRUE), 
    allow_retry = TRUE, 
    random_answer_order = TRUE, 
    type = "multiple"
)
```

```{r q10}
  question("Assumptions for Linear Regression are ",
    answer("Linearity", correct = TRUE),
    answer("Homoscedasticity", correct = TRUE),
    answer("No auto-correlation", correct = TRUE),
    answer("Normality", correct = TRUE), 
    allow_retry = TRUE, 
    random_answer_order = TRUE, 
    type = "multiple"
)
```

#### True or False

```{r q11}
  question("Logestic Regression is used when we try to predict binary value",
    answer("TRUE", correct = TRUE),
    answer("FALSE"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```


```{r q12}
  question("Logestic Regression is used when we try to predict binary value",
    answer("TRUE", correct = TRUE),
    answer("FALSE"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```


```{r q13}
  question("logit transofrmation is to convert values between -Inf and +Inf to values between 0 and 1",
    answer("TRUE", correct = TRUE),
    answer("FALSE"),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```


```{r q14}
  question("The **bias** is an error from sensitivity to small fluctuations in the training set.",
    answer("TRUE"),
    answer("FALSE", correct = TRUE),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```

```{r q15}
  question("The **variance** is an error from erroneous assumptions in the learning algorithm",
    answer("TRUE"),
    answer("FALSE", correct = TRUE),
    random_answer_order = TRUE,
    allow_retry = TRUE
)
```


## K nearest neighborhood 

### Exercise

```{r preperations}
library("tibble")
df_knn <- tribble(
~Object,   ~Gender, ~Party,	       ~Age,	~Weight,	~Income,	~Class,
    "A",	"Female",	"Independent",	20,	    120,	   40000,	   "Yes",
    "B",	"Male",   "Republican",	  60,	    130,	   80000,	   "No",
    "C",	"Male",   "Democrat",  	  35,	    150,	   50000,	   "Yes",
    "D",	"Female",	"Democrat",  	  45,	    160,	   30000,	   "Yes",
    "E",	"Female",	"Republican",   30,   	145,	   32000,	   "No"
) %>% mutate_at(c("Gender","Party", "Class"),factor)
knn_new <- tribble(
~Object,   ~Gender, ~Party,	       ~Age,	~Weight,	~Income,	~Class,
    "W",	"Female",	"Independent",	20,	    120,	   40000,   "Yes",
    "X",	"Male",	  "Democrat",	    50,	    134,	   44000,	   "No"
) %>% mutate_at(c("Gender","Party", "Class"),factor)
```

A researcher sent you and email with a traing set `df_knn` and test set `knn_new`. He explained over the phone that he would like to build a model that tells if the person is lying or being honest. Now your job is to use the data and build classification model using KNN. 

<div id="q16-hint">
**Hint:** :  You might want to normalize quantative variables.
</div>


```{r q16, exercise = TRUE, exercise.lines = 30,exercise.setup = "preperations", exercise.eval = FALSE}
#raw data
df_knn

#the new data
knn_new
```




```{r q16-solution}
#raw data
combined_df <-rbind(df_knn,knn_new)
#preprocessing data and scaling them 
mat_formula <- Class~Gender +Party+scale(Age)+scale(Weight)+scale(Income)-1

comb_df_mat <-model.matrix( mat_formula, data = combined_df)

dist_df<-dist(comb_df_mat,upper = TRUE, diag = TRUE)
dist_df<-as.matrix(dist_df)

sort(dist_df[6,],decreasing = FALSE) # YES if k = 3 we have 1:"YES" 5:"NO" 3:"YES"
sort(dist_df[7,],decreasing = FALSE) # YES if k = 3 we have 3:"YES" 4:"YES" 5:"NO"  

#one other solutions 
library("caret")
dummy_m <- dummyVars(Class ~ ., data = df_knn[,-1])
preproces_m <- preProcess(df_knn[,-1], method = c("center", "scale"))


tr_set <- predict(object = preproces_m, newdata = df_knn[,-1]) 
tr_set <- predict(object = dummy_m, newdata = tr_set) %>% data.frame()
tr_set$Class <- df_knn$Class
knn_m<-train(Class ~ ., data = tr_set, method = "knn", tuneGrid = expand.grid(k = 3), tuneLength = 1, trControl = trainControl(method = "none", number = 1))

te_set <- predict(object = preproces_m, newdata = knn_new[,-1]) 
te_set <- predict(object = dummy_m, newdata = knn_new) %>% data.frame()
te_set$Class <- knn_new$Class


predict(knn_m, newdata = te_set)

```




















